---
title: "Inferential deep learning trials on KickCap dataset: statistical analyses"
author: "Joël Singer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Learning analyses

Learning curves for small (<1% of dataset) training set, from one situation of one participant.

```{r learning curves, echo=FALSE, warning=FALSE, message=FALSE}
# Load data
training_p2p <- read_csv("results/learning_results/Learning_results_original.csv")

training_seq2seq <- read_csv("results/learning_results/Learning_results_simple_seq2seq.csv")

training_soleformer <- read_csv("results/learning_results/Learning_results_soleformer.csv")

# Plot train and val losses for each model in function of epoch
lc_p2p <- training_p2p |> 
  ggplot(aes(x = epoch)) +
  geom_line(aes(y = train_loss, color = "train_loss")) +
  geom_line(aes(y = val_loss, color = "val_loss")) +
  labs(title = "P2P", x = "Epoch", y = "Loss") +
  scale_color_manual(values = c("train_loss" = "blue", "val_loss" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank())

lc_seq2seq <- training_seq2seq |>
  ggplot(aes(x = epoch)) +
  geom_line(aes(y = train_loss, color = "train_loss")) +
  geom_line(aes(y = val_loss, color = "val_loss")) +
  labs(title = "Seq2Seq", x = "Epoch", y = "Loss") +
  scale_color_manual(values = c("train_loss" = "blue", "val_loss" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank())

lc_soleformer <- training_soleformer |>
  ggplot(aes(x = epoch)) +
  geom_line(aes(y = train_loss, color = "train_loss")) +
  geom_line(aes(y = val_loss, color = "val_loss")) +
  labs(title = "SoleFormer", x = "Epoch", y = "Loss") +
  scale_color_manual(values = c("train_loss" = "blue", "val_loss" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank())

lc_p2p
lc_seq2seq
lc_soleformer

```

The learning of P2P-Insole and its sequence-to-sequence variant is very poor. On the other hand, SoleFormer shows a standard learning curve, also showing that 40 epochs are not enough for an optimal learning.

# MPJPE

Mean Per Joint Position Error (MPJPE) is an intuitive measure of correspondance between a ground-truth and a prediction. It is the mean euclidean distance between the predicted and the ground-truth 3D joint positions, averaged across all joints and all frames. The lower the MPJPE, the better the model's performance in predicting accurate 3D poses.

```{r mpjpe, echo=FALSE, warning=FALSE, message=FALSE}

```

















Discussion

We can obviously observe that the P2P-Insole and its sequence-to-sequence version are not learning as much as we would expect. It relates to their animations, where we could see that the model is not able to learn the dynamic mapping between feet data and whole-body motion (even with the full dataset). The model is thus stuck into always predicting an average pose, whatever the input. The SoleFormer, on the other hand, is showing standard expected learning curves. However, when trained with the full dataset or even with only the data of one participant (not shown), it outputs similarly as the two former. 

The training data provided in this repository corresponds to roughly 40 seconds of complex motion: shadow-boxing: the participant is free to simulate a fight in the air. The algorithm is still capable to learn from that and infer quite well 10 seconds at the other end of the same shadow-boxing situation - considering the training dataset size and the very poor results of the other models. The motion complexity seems to be the major issue encountered with our dataset, compared to the literature. Indeed, the movements previously investigated corresponded to daily-life situations or sports where movements are quite similar to each other or where feet motion is easily relatable to the whole movement (golf, skiing, table tennis). In lieu, the kickboxing activity is characterized by a large variety of movements, that often exhibit for a same movement different feet patterns, or for a same feet pattern different whole-body movements. 

In fact, this project recently uncovered the question about what is a kickboxing movement, more specifically: can we classify kickboxing movements? Indeed, among our current perspectives, we are planning to adjust the weights applied to features at the input of the model in function of the movement performed. Through this method, we hope to optimise the inference by emphasizing the power of the most descriptive features of each technique (previously defined by PCA or machine learning classifiers). However, a kickboxing movement is always comprised in a bigger context: a combination, a displacement, a shock, a defense, etc. The movement is often "beginning" - "emerging" would be a better word - while another movement is being performed. It is thus difficult and maybe incorrect to clearly distinguish one movement from another in a sequence of frames. Uncertainty theory could be a good way to deal with this issue, by allowing a "soft" classification of movements.

Going back to our results, we can see that SoleFormer, that has a more elegant architecture, is able to learn more. Although its number of parameters is less than half the number of the other models. This is a good example of how the architecture of a model can be more important than its size, and making its architecture suits the data in input as well as the wanted output is actually part of what we are investigating through those literature models' trials. Our final objective is to design from scratch a model specific to kickboxing and the measurement material we are using. The double cycle-loss architecture, adding two loss functions differently handling pressure and IMU data, is precisely one way to make the model more elegant. Another aspect certainly allowing the model to learn a dynamic mapping is the encoding block at the pressure features input: a GNN. Because the pressure sensors are spatially distributed on the foot, they can be represented as a graph from which the GNN allow to efficiently encode the relations between sensors (e. g., opposition of the medial and lateral parts). The last big improvement of this model is the two-stream cross-attention mechanism, that allow the transformer block to link IMU and pressure data: rolling a foot from on side to another describes a specific pressure pattern that is also distinguishable in the roll gyroscopic data. We did not yet realised the ablation study on SoleFormer, but it will give us the best architecture elements for our problem.

