---
title: "Inferential deep learning trials on KickCap dataset: statistical analyses"
author: "Joël Singer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(ggplot2)
library(reshape2)
library(stringr)
library(gridExtra)
library(grid)
library(gtable)
```

# Learning analyses

Learning curves for small (<1% of dataset) training set, from one situation of one participant.

```{r learning curves, echo=FALSE, warning=FALSE, message=FALSE}
# Load data
training_p2p <- read_csv("results/learning_results/Learning_results_original.csv")

training_seq2seq <- read_csv("results/learning_results/Learning_results_simple_seq2seq.csv")

training_soleformer <- read_csv("results/learning_results/Learning_results_soleformer.csv")

# Plot train and val losses for each model in function of epoch
lc_p2p <- training_p2p |> 
  ggplot(aes(x = epoch)) +
  geom_line(aes(y = train_loss, color = "train_loss")) +
  geom_line(aes(y = val_loss, color = "val_loss")) +
  labs(title = "P2P", x = "Epoch", y = "Loss") +
  scale_color_manual(values = c("train_loss" = "blue", "val_loss" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank())

lc_seq2seq <- training_seq2seq |>
  ggplot(aes(x = epoch)) +
  geom_line(aes(y = train_loss, color = "train_loss")) +
  geom_line(aes(y = val_loss, color = "val_loss")) +
  labs(title = "Seq2Seq", x = "Epoch", y = "Loss") +
  scale_color_manual(values = c("train_loss" = "blue", "val_loss" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank())

lc_soleformer <- training_soleformer |>
  ggplot(aes(x = epoch)) +
  geom_line(aes(y = train_loss, color = "train_loss")) +
  geom_line(aes(y = val_loss, color = "val_loss")) +
  labs(title = "SoleFormer", x = "Epoch", y = "Loss") +
  scale_color_manual(values = c("train_loss" = "blue", "val_loss" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank())

lc_p2p
lc_seq2seq
lc_soleformer

```

The learning of P2P-Insole and its sequence-to-sequence variant is very poor. On the other hand, SoleFormer shows a standard learning curve, also showing that 40 epochs are not enough for an optimal learning.

# Prediction analyses

## Initialization

```{r analyses init, echo=FALSE, warning=FALSE, message=FALSE}
# Load data
gt_files <- c(
  "data/test_data/skeleton/Awinda_001CcSs_3_shadow_test.csv",
  "data/test_data/skeleton/Awinda_001CcSs_3_shadow_test.csv",
  "data/clean_data/Awinda_targets_soleformer/AwindaTarget_001CcSs_3_shadow_test.csv"
)

pred_files <- c(
  "results/output/Predicted_skeleton_001CcSs_3_shadow_test_original.csv",
  "results/output/Predicted_skeleton_001CcSs_3_shadow_test_simple_seq2seq.csv",
  "results/output/Predicted_skeleton_001CcSs_3_shadow_test_soleformer.csv"
)

# Names of algo for titles
method_names <- c(
  "P2P-Insole",
  "P2P-Insole seq2seq",
  "SoleFormer"
)

# Joint definitions
joint_names <- c(
  "Pelvis","L5","L3","T12","T8","Neck","Head",
  "Right Shoulder","Right Upper Arm","Right Forearm","Right Hand",
  "Left Shoulder","Left Upper Arm","Left Forearm","Left Hand",
  "Right Upper Leg","Right Lower Leg","Right Foot","Right Toe",
  "Left Upper Leg","Left Lower Leg","Left Foot","Left Toe"
)

upper_idx <- 8:15
lower_idx <- 16:23
trunk_idx <- 1:7


# Frame alignment function
align_dataframes <- function(gt, pred) {

  has_frame_gt <- "Frame" %in% colnames(gt)
  has_frame_pred <- "Frame" %in% colnames(pred)

  # Case 1: both have Frame -> align by intersection
  if (has_frame_gt && has_frame_pred) {

    common_frames <- intersect(gt$Frame, pred$Frame)

    gt <- gt[Frame %in% common_frames]
    pred <- pred[Frame %in% common_frames]

    # ensure same order
    setorder(gt, Frame)
    setorder(pred, Frame)

    return(list(gt=gt, pred=pred))
  }

  # Case 2: no Frame -> fallback to truncation
  n <- min(nrow(gt), nrow(pred))

  gt <- gt[1:n]
  pred <- pred[1:n]

  list(gt=gt, pred=pred)
}


# Col parser handling both formats and ignoring extra cols
extract_positions <- function(df) {
  cols <- colnames(df)

  # Case 1: X.0, Y.0...
  if (any(grepl("^X\\.0$", cols))) {
    X <- df[, grep("^X\\.", cols), with=FALSE]
    Y <- df[, grep("^Y\\.", cols), with=FALSE]
    Z <- df[, grep("^Z\\.", cols), with=FALSE]

  } else {
    # Case 2: pos::Joint x
    pos_cols <- grep("^pos::", cols, value=TRUE)

    X <- df[, pos_cols[grepl(" x$", pos_cols)], with=FALSE]
    Y <- df[, pos_cols[grepl(" y$", pos_cols)], with=FALSE]
    Z <- df[, pos_cols[grepl(" z$", pos_cols)], with=FALSE]
  }

  list(X=as.matrix(X), Y=as.matrix(Y), Z=as.matrix(Z))
}


# Shared preparation for each metric
prepare_pair <- function(gt_path, pred_path) {

  gt <- fread(gt_path)
  pred <- fread(pred_path)

  aligned <- align_dataframes(gt, pred)

  # Positions
  gt_pos <- extract_positions(aligned$gt)
  pred_pos <- extract_positions(aligned$pred)

  # Angles (if present)
  has_angles <- any(grepl("^ang::", colnames(pred)))

  if (has_angles) {
    ang <- extract_angles(aligned$gt, aligned$pred)
  } else {
    ang <- NULL
  }

  list(
    gt_pos = gt_pos,
    pred_pos = pred_pos,
    angles = ang
  )
}
```

```{r ang helper functions, echo=FALSE, warning=FALSE, message=FALSE}
# Different col name handling
normalize_names <- function(x) {
  x <- trimws(x)
  x <- gsub("\\s+", " ", x)
  x <- tolower(x)
  x
}

# Angle extractor for the SoleFormer output
extract_angles <- function(gt, pred) {

  gt_cols <- grep("^ang::", colnames(gt), value=TRUE)
  pred_cols <- grep("^ang::", colnames(pred), value=TRUE)

  # REMOVE WRIST COLUMNS (case-insensitive)
  gt_cols <- gt_cols[!grepl("wrist", gt_cols, ignore.case=TRUE)]
  pred_cols <- pred_cols[!grepl("wrist", pred_cols, ignore.case=TRUE)]

  gt_norm <- normalize_names(gt_cols)
  pred_norm <- normalize_names(pred_cols)

  common_norm <- intersect(gt_norm, pred_norm)

  if (length(common_norm) == 0) {
    stop("No common angle columns found after filtering and normalization")
  }

  gt_match_idx <- match(common_norm, gt_norm)
  pred_match_idx <- match(common_norm, pred_norm)

  gt_sel <- gt_cols[gt_match_idx]
  pred_sel <- pred_cols[pred_match_idx]

  gt_mat <- as.matrix(gt[, ..gt_sel])
  pred_mat <- as.matrix(pred[, ..pred_sel])

  list(
    gt = gt_mat,
    pred = pred_mat,
    names = gt_sel
  )
}

# Name cleaner
clean_angle_names <- function(names) {
  names <- gsub("^ang::", "", names)
  names <- gsub(" +", " ", names)
  names
}
```


```{r body parts helper functions, echo=FALSE, warning=FALSE, message=FALSE}
get_segments <- function(n_joints) {
  list(
    full  = 1:n_joints,
    upper = upper_idx,
    lower = lower_idx,
    trunk = trunk_idx
  )
}

aggregate_segments <- function(matrix, segments, fun = mean) {

  sapply(segments, function(idx) {
    fun(matrix[, idx], na.rm=TRUE)
  })
}
```

## Plot functions

```{r heatmap, echo=FALSE, warning=FALSE, message=FALSE}
# Heatmap
plot_heatmap <- function(dist_matrix, title, mpjpe, sd_val, metric_name, unit, fill_label) {
  df <- melt(dist_matrix)
  colnames(df) <- c("Frame","Joint","Error")

  df$Joint <- factor(df$Joint, labels=joint_names)

  ggplot(df, aes(Frame, Joint, fill=Error)) +
    geom_tile() +
    scale_fill_gradient(low="green", high="red", name=fill_label) +
    ggtitle(title) +
    annotate("text",
             x = Inf, y = Inf,
             label = sprintf("%s: %.2f ± %.2f %s", metric_name, mpjpe, sd_val, unit),
             hjust = 1.1, vjust = 1.5) +
    theme_minimal()
}
```

```{r ranking joints, echo=FALSE, warning=FALSE, message=FALSE}
# Histogram of mean position error for each joint
plot_joint_ranking_metric <- function(values, title,
                              y_label = "Metric (unit)") {

  df <- data.frame(
    Joint = joint_names,
    Value = values
  )
  
  # sort descendingly
  df <- df[order(-df$Value), ]
  df$Joint <- factor(df$Joint, levels=df$Joint)

  ggplot(df, aes(x=Joint, y=Value)) +
    geom_bar(stat="identity", fill="steelblue") +
    coord_flip() +
    ggtitle(title) +
    ylab(y_label) +
    theme_minimal()
}
```


```{r histogram, echo=FALSE, warning=FALSE, message=FALSE}
# Histogram for body parts
plot_segment_metric <- function(values, title,
                                y_label = "Metric (unit)") {

  df <- data.frame(
    Segment = c("Full Body", "Upper Limb", "Lower Limb", "Trunk"),
    Value = values
  )

  ggplot(df, aes(x=Segment, y=Value, fill=Segment)) +
    geom_bar(stat="identity", width=0.6) +
    ggtitle(title) +
    ylab(y_label) +
    theme_minimal() +
    theme(legend.position="none")
}
```


## Frame-wise pose estimation metrics

These metrics are founded on differences frame-by-frame and joint-by-joint between the predicted and the ground-truth poses. They are the most intuitive and commonly used metrics in pose estimation, as they directly measure the accuracy of the predicted joint positions compared to the ground-truth. They can be computed for each joint and each frame, allowing for detailed analyses of where and when the model performs well or poorly.



### MPJPE

Mean Per Joint Position Error (MPJPE) is an intuitive measure of correspondance between a ground-truth and a prediction. It is the mean euclidean distance between the predicted and the ground-truth 3D joint positions, averaged across all joints and all frames. The lower the MPJPE, the better the model's performance in predicting accurate 3D poses.

```{r mpjpe comput, echo=FALSE, warning=FALSE, message=FALSE}
# MPJPE
compute_mpjpe <- function(gt, pred) {
  dx <- pred$X - gt$X
  dy <- pred$Y - gt$Y
  dz <- pred$Z - gt$Z

  dist <- sqrt(dx^2 + dy^2 + dz^2)

  list(
    per_joint = colMeans(dist, na.rm=TRUE),
    per_frame = rowMeans(dist, na.rm=TRUE),
    full = mean(dist, na.rm=TRUE),
    sd = sd(dist, na.rm=TRUE),
    matrix = dist
  )
}
```


```{r mpjpe main loop, echo=FALSE, warning=FALSE, message=FALSE}
# Main MPJPE processing loop
for (i in seq_along(gt_files)) {

  data <- prepare_pair(gt_files[i], pred_files[i])
  method <- method_names[i]

  res <- compute_mpjpe(data$gt_pos, data$pred_pos)

  print(plot_heatmap(res$matrix,
                    paste("Temporal Position Error Heatmap -", method),
                    res$full, res$sd,
                    metric_name = "MPJPE",
                    unit = "m",
                    fill_label = "Error (m)"))

  segments <- get_segments(ncol(res$matrix))
  res_segments <- aggregate_segments(res$matrix, segments)

  print(plot_segment_metric(res_segments,
                            paste("Body Part MPJPE -", method),
                            y_label = "MPJPE (m)"))

  print(plot_joint_ranking_metric(res$per_joint,
                          paste("Joint 3D Position Error Ranking -", ... = method),
                          y_label = "Mean Position Error (m)"))
}
```

### PA-MPJPE
Procrustes Aligned Mean Per Joint Position Error (PA-MPJPE) is a measure of the average distance between the predicted and ground-truth joint positions after aligning them using Procrustes analysis. This alignment removes differences in scale, rotation, and translation, allowing for a more accurate assessment of the pose estimation performance. The lower the PA-MPJPE, the better the model's performance in predicting accurate 3D poses.

```{r pa-mpjpe comput, echo=FALSE, warning=FALSE, message=FALSE}
# PA-MPJPE computation
compute_pa_mpjpe <- function(gt, pred) {

  n_frames <- nrow(gt$X)
  n_joints <- ncol(gt$X)

  dist <- matrix(NA, nrow=n_frames, ncol=n_joints)

  for (f in 1:n_frames) {

    # build 3D matrices (J x 3)
    G <- cbind(gt$X[f,], gt$Y[f,], gt$Z[f,])
    P <- cbind(pred$X[f,], pred$Y[f,], pred$Z[f,])

    # center
    Gc <- scale(G, center=TRUE, scale=FALSE)
    Pc <- scale(P, center=TRUE, scale=FALSE)

    # SVD for optimal rotation
    H <- t(Pc) %*% Gc
    svd_res <- svd(H)
    R <- svd_res$v %*% t(svd_res$u)

    # reflection correction
    if (det(R) < 0) {
      svd_res$v[,3] <- -svd_res$v[,3]
      R <- svd_res$v %*% t(svd_res$u)
    }

    # align
    P_aligned <- Pc %*% R

    # compute distances
    d <- sqrt(rowSums((P_aligned - Gc)^2))
    dist[f, ] <- d
  }

  list(
    per_joint = colMeans(dist, na.rm=TRUE),
    per_frame = rowMeans(dist, na.rm=TRUE),
    full = mean(dist, na.rm=TRUE),
    sd = sd(dist, na.rm=TRUE),
    matrix = dist
  )
}
```


```{r pa-mpjpe main loop, echo=FALSE, warning=FALSE, message=FALSE}
# Main processing loop for PA-MPJPE
for (i in seq_along(gt_files)) {

  data <- prepare_pair(gt_files[i], pred_files[i])
  method <- method_names[i]

  res <- compute_pa_mpjpe(data$gt_pos, data$pred_pos)

  print(plot_heatmap(res$matrix,
                     paste("Temporal Procrustes-Aligned Position Error Heatmap -", method),
                     res$full, res$sd,
                     fill_label = "Error (m)",
                     metric_name = "PA-MPJPE",
                     unit = "m"))

  segments <- get_segments(ncol(res$matrix))
  res_segments <- aggregate_segments(res$matrix, segments)

  print(plot_segment_metric(res_segments,
                            paste("Body Part PA-MPJPE -", method),
                            y_label = "PA-MPJPE (m)"))

  print(plot_joint_ranking_metric(res$per_joint,
                          paste("Joint PA-JPE Ranking -", method),
                          y_label = "Mean Procrustes-Aligned Position Error (m)"))
}
```

### MAJAE
Mean Absolute Joint Angle Error (MAJAE) is a measure of the average absolute difference in joint angles between the predicted and ground-truth poses. It is calculated by directly computing the difference between SoleFormer input and output joint angles, that are already taken from the Awinda data and computed by the SoleFormer pipeline.

```{r majae comp, echo=FALSE, warning=FALSE, message=FALSE}
# MAJAE computation
compute_majae <- function(gt_ang, pred_ang) {
  diff <- abs(pred_ang - gt_ang)

  list(
    mean = mean(diff, na.rm=TRUE),
    sd = sd(diff, na.rm=TRUE),
    matrix = diff
  )
}
```


```{r joints ang ranking, echo=FALSE, warning=FALSE, message=FALSE}
# Histogram of mean angle error for each joint
plot_angle_ranking <- function(angle_matrix, angle_names, title) {

  angle_majae <- colMeans(angle_matrix, na.rm=TRUE)

  df <- data.frame(
    Angle = angle_names,
    MAJAE = angle_majae
  )

  df <- df[order(-df$MAJAE), ]
  df$Angle <- factor(df$Angle, levels=df$Angle)

  ggplot(df, aes(x=Angle, y=MAJAE)) +
    geom_bar(stat="identity", fill="steelblue") +
    coord_flip() +
    ggtitle(title) +
    ylab("Mean Angle Error (degrees)") +
    xlab("Joint Angle") +
    theme_minimal()
}
```


```{r body segments ang, echo=FALSE, warning=FALSE, message=FALSE}
# Group angles by body part based on their names
get_angle_groups <- function(angle_names) {

  upper <- grepl("Shoulder|Elbow", angle_names, ignore.case=TRUE)
  lower <- grepl("Hip|Knee|Ankle|Foot", angle_names, ignore.case=TRUE)
  trunk <- grepl("L5S1|L4L3|L1T12|T9T8|T1C7|C1 Head", angle_names, ignore.case=TRUE)

  list(
    upper = upper,
    lower = lower,
    trunk = trunk,
    full  = rep(TRUE, length(angle_names))
  )
}

# Plot MAJAE for body segments
plot_angle_segments <- function(angle_matrix, angle_names, title) {

  groups <- get_angle_groups(angle_names)

  majae_full  <- mean(angle_matrix[, groups$full],  na.rm=TRUE)
  majae_upper <- mean(angle_matrix[, groups$upper], na.rm=TRUE)
  majae_lower <- mean(angle_matrix[, groups$lower], na.rm=TRUE)
  majae_trunk <- mean(angle_matrix[, groups$trunk], na.rm=TRUE)

  df <- data.frame(
    Segment = c("Full Body", "Upper Limb", "Lower Limb", "Trunk"),
    MAJAE = c(majae_full, majae_upper, majae_lower, majae_trunk)
  )

  ggplot(df, aes(x=Segment, y=MAJAE, fill=Segment)) +
    geom_bar(stat="identity", width=0.6) +
    ggtitle(title) +
    ylab("MAJAE (degrees)") +
    theme_minimal() +
    theme(legend.position="none")
}
```


```{r ang main loop, echo=FALSE, warning=FALSE, message=FALSE}
# Main MAJAE processing loop
for (i in seq_along(gt_files)) {

  data <- prepare_pair(gt_files[i], pred_files[i])
  method <- method_names[i]

  if (!is.null(data$angles)) {

    ang <- data$angles

    res_ang <- compute_majae(ang$gt, ang$pred)

    # Histogram (distribution)
    df <- melt(res_ang$matrix)

    p <- ggplot(df, aes(value)) +
      geom_histogram(bins=50, fill="blue", alpha=0.6) +
      ggtitle(paste("Angle Error Distribution -", method)) +
      annotate("text",
               x = Inf, y = Inf,
               label = sprintf("MAJAE: %.2f ± %.2f degrees",
                               res_ang$mean, res_ang$sd),
               hjust = 1.1, vjust = 1.5) +
      theme_minimal()

    print(p)

    # Ranking
    print(plot_angle_ranking(res_ang$matrix,
                             clean_angle_names(ang$names),
                             paste("Joint Angle Error Ranking -", method)))

    # Segment-level (same philosophy as MPJPE)
    print(plot_angle_segments(res_ang$matrix,
                              clean_angle_names(ang$names),
                              paste("Body Part MAJAE -", method)))
  }
}
```

### MPJVE
Mean Per Joint Velocity Error (MPJVE) is a measure of smoothness and temporal consistency in pose estimation. It is calculated by first computing the velocity of each joint in both the predicted and ground-truth poses (i.e., the change in joint position between consecutive frames) and then taking the mean of the absolute differences in these velocities across all joints and frames. A lower MPJVE indicates that the model is better at capturing the dynamic aspects of motion, producing smoother and more realistic pose estimations over time.

```{r velocity, echo=FALSE, warning=FALSE, message=FALSE}
# Velocity computation
compute_velocity <- function(pos, fps = 60) {

  list(
    X = (pos$X[-1, ] - pos$X[-nrow(pos$X), ]) * fps,
    Y = (pos$Y[-1, ] - pos$Y[-nrow(pos$Y), ]) * fps,
    Z = (pos$Z[-1, ] - pos$Z[-nrow(pos$Z), ]) * fps
  )
}
```


```{r mpjve, echo=FALSE, warning=FALSE, message=FALSE}

compute_mpjve <- function(gt, pred) {

  gt_v <- compute_velocity(gt)
  pred_v <- compute_velocity(pred)

  compute_mpjpe(gt_v, pred_v)
}
```


```{r main mpjve loop, echo=FALSE, warning=FALSE, message=FALSE}
# Main processing loop for MPJVE
for (i in seq_along(gt_files)) {

  data <- prepare_pair(gt_files[i], pred_files[i])
  method <- method_names[i]

  res <- compute_mpjve(data$gt_pos, data$pred_pos)

  print(plot_heatmap(res$matrix,
                     paste("Temporal Velocity Error Heatmap -", method),
                     res$full, res$sd,
                     fill_label = "Error (m/s)",
                     metric_name = "MPJVE",
                     unit = "m/s"))

  segments <- get_segments(ncol(res$matrix))
  res_segments <- aggregate_segments(res$matrix, segments)

  print(plot_segment_metric(res_segments,
                            paste("Body Part MPJVE -", method),
                            y_label = "MPJVE (m/s)"))

  print(plot_joint_ranking_metric(res$per_joint,
                          paste("Joint 3D Velocity Error Ranking -", method),
                          y_label = "Mean Velocity Error (m/s)"))
}
```




## Pose estimation realism metrics

These metrics are founded on the analysis of the predicted and ground-truth motions as a whole, rather than on frame-by-frame differences. They capture the overall quality of the motion, its "realism". Indeed, static metrics like MPJPE can be low even if the predicted motion is not realistic, for example if the model is just predicting an average pose. On the other hand, metrics like velocity correlation and motion energy ratio can capture whether the predicted motion has similar dynamics and energy to the ground-truth. This is crucial to distinguish algorithms not learning dynamically from those that can have bad static results (because of the complexity of the movement) but that are learning a dynamic mapping and thus can produce more realistic motions.


### Motion Energy Ratio
Motion Energy Ratio (MER) is a measure of how well the predicted motion captures the overall energy of the ground-truth motion: meaning the amplitude of movements. As for velocity correlation, it focuses on the motion itself instead of mere differences. 

```{r mot ener comput, echo=FALSE, warning=FALSE, message=FALSE}
# Motion energy computation
compute_motion_energy <- function(pos) {

  v <- compute_velocity(pos)

  # Per-joint energy
  energy_per_joint <- colMeans(
    v$X^2 + v$Y^2 + v$Z^2,
    na.rm = TRUE
  )

  # Full-body energy (scalar)
  energy_full <- mean(energy_per_joint, na.rm=TRUE)

  list(
    per_joint = energy_per_joint,
    full = energy_full
  )
}

# Energy ratio per joint
compute_motion_energy_ratio <- function(gt, pred) {

  e_gt <- compute_motion_energy(gt)
  e_pred <- compute_motion_energy(pred)

  ratio_per_joint <- e_pred$per_joint / e_gt$per_joint
  ratio_per_joint[!is.finite(ratio_per_joint)] <- NA

  ratio_full <- e_pred$full / e_gt$full

  list(
    gt = e_gt,
    pred = e_pred,
    ratio = list(
      per_joint = ratio_per_joint,
      full = ratio_full
    )
  )
}


```

```{r mot ener plot, echo=FALSE, warning=FALSE, message=FALSE}
# Motion energy plot
plot_motion_energy <- function(res, title) {

  df <- data.frame(
    Type = c("Ground Truth", "Prediction"),
    Energy = c(res$gt$full, res$pred$full)
  )

  ggplot(df, aes(x=Type, y=Energy, fill=Type)) +
    geom_bar(stat="identity", width=0.6) +
    ggtitle(paste(title, sprintf("(Ratio = %.2f)", res$ratio$full))) +
    ylab("Motion Energy (m²/s²)") +
    theme_minimal() +
    theme(legend.position="none")
}



# Aggregation of energy by body segments
compute_energy_segments <- function(res) {

  segments <- get_segments(length(res$gt$per_joint))

  gt_seg <- sapply(segments, function(idx)
    mean(res$gt$per_joint[idx], na.rm=TRUE))

  pred_seg <- sapply(segments, function(idx)
    mean(res$pred$per_joint[idx], na.rm=TRUE))

  ratio_seg <- sapply(segments, function(idx)
    mean(res$ratio$per_joint[idx], na.rm=TRUE))

  list(
    gt = gt_seg,
    pred = pred_seg,
    ratio = ratio_seg
  )
}
```

```{r mot ener main loop, echo=FALSE, warning=FALSE, message=FALSE}
# Main motion energy loop
for (i in seq_along(gt_files)) {

  data <- prepare_pair(gt_files[i], pred_files[i])
  method <- method_names[i]

  res_energy <- compute_motion_energy_ratio(data$gt_pos, data$pred_pos)

  # Plot (full-body only)
  print(plot_motion_energy(res_energy,
                          paste("Motion Energy -", method)))

  # Segment values (for table later)
  res_energy_seg <- compute_energy_segments(res_energy)

  # -> store this in summary_df later
}
```


### Velocity Correlation
Velocity Correlation (VC) is a measure of how well the predicted joint velocities match the ground-truth joint velocities. It goes one step further than computing the raw differences in velocity (MPJVE) by correlating time-series. It captures the temporal dynamics and patterns of movement rather than just the magnitude of errors.

```{r v corr comput, echo=FALSE, warning=FALSE, message=FALSE}
# Velocity correlation computation
compute_velocity_correlation <- function(gt, pred) {

  gt_v <- compute_velocity(gt)
  pred_v <- compute_velocity(pred)

  n_joints <- ncol(gt_v$X)
  corrs <- rep(NA, n_joints)

  for (j in 1:n_joints) {

    v_gt <- c(gt_v$X[,j], gt_v$Y[,j], gt_v$Z[,j])
    v_pr <- c(pred_v$X[,j], pred_v$Y[,j], pred_v$Z[,j])

    if (sd(v_gt) > 0 && sd(v_pr) > 0) {
      corrs[j] <- cor(v_gt, v_pr)
    }
  }

  list(
    per_joint = corrs,
    mean = mean(corrs, na.rm=TRUE)
  )
}
```

```{r v corr plot, echo=FALSE, warning=FALSE, message=FALSE}
# Velocity correlation plot
plot_velocity_correlation <- function(corrs, title) {

  df <- data.frame(
    Joint = joint_names,
    Correlation = corrs
  )

  df <- df[order(-df$Correlation), ]
  df$Joint <- factor(df$Joint, levels=df$Joint)

  ggplot(df, aes(x=Joint, y=Correlation)) +
    geom_bar(stat="identity", fill="darkgreen") +
    coord_flip() +
    ggtitle(title) +
    ylab("Velocity Correlation") +
    ylim(-1, 1) +
    theme_minimal()
}
```

```{r v corr main loop, echo=FALSE, warning=FALSE, message=FALSE}
# Main velocity correlation loop
for (i in seq_along(gt_files)) {

  data <- prepare_pair(gt_files[i], pred_files[i])
  method <- method_names[i]

  res_corr <- compute_velocity_correlation(data$gt_pos, data$pred_pos)

  print(plot_velocity_correlation(res_corr$per_joint,
                                  paste("Velocity Correlation -", method)))

  cat("Mean velocity correlation:", res_corr$mean, "\n")
}
```


## Summary tables

These tables displays the 6 metrics for each model and for each body part, allowing for a quick comparison of their performance across different aspects of pose estimation quality, and for the identification of difficult body segments to be predicted.

### Console display

```{r summary table, echo=FALSE, warning=FALSE, message=FALSE}
# Precompute energy segments for all methods to use in the summary table
energy_segments_list <- list()

for (i in seq_along(gt_files)) {

  data <- prepare_pair(gt_files[i], pred_files[i])
  method <- method_names[i]

  res_energy <- compute_motion_energy_ratio(data$gt_pos, data$pred_pos)
  res_energy_seg <- compute_energy_segments(res_energy)

  energy_segments_list[[method]] <- res_energy_seg$ratio
}

# Table builder
build_summary_table <- function(gt_files, pred_files, method_names, energy_segments_list) {

  results_list <- list()

  for (i in seq_along(gt_files)) {

    data <- prepare_pair(gt_files[i], pred_files[i])
    method <- method_names[i]

    seg <- get_segments(ncol(data$gt_pos$X))

    # --- Metrics ---
    mpjpe  <- compute_mpjpe(data$gt_pos, data$pred_pos)
    pa     <- compute_pa_mpjpe(data$gt_pos, data$pred_pos)
    mpjve  <- compute_mpjve(data$gt_pos, data$pred_pos)
    corr   <- compute_velocity_correlation(data$gt_pos, data$pred_pos)

    # --- Segment aggregation ---
    mpjpe_s <- aggregate_segments(mpjpe$matrix, seg)
    pa_s    <- aggregate_segments(pa$matrix, seg)
    mpjve_s <- aggregate_segments(mpjve$matrix, seg)

    # correlation per segment
    corr_s <- sapply(seg, function(idx) {
      mean(corr$per_joint[idx], na.rm=TRUE)
    })

    # energy ratio (per segment)
    energy_s <- energy_segments_list[[method]]

    # --- Angles (if available) ---
    if (!is.null(data$angles)) {

      ang <- data$angles
      majae <- compute_majae(ang$gt, ang$pred)

      angle_names <- clean_angle_names(ang$names)
      groups <- get_angle_groups(angle_names)

      majae_s <- c(
        full  = mean(majae$matrix, na.rm=TRUE),
        upper = mean(majae$matrix[, groups$upper], na.rm=TRUE),
        lower = mean(majae$matrix[, groups$lower], na.rm=TRUE),
        trunk = mean(majae$matrix[, groups$trunk], na.rm=TRUE)
      )

    } else {
      majae_s <- rep(NA, 4)
      names(majae_s) <- names(seg)
    }

    # --- Combine into one row ---
    row <- c()

    for (part in names(seg)) {

      row <- c(row,
               mpjpe_s[part],
               pa_s[part],
               mpjve_s[part],
               corr_s[part],
               energy_s[part],
               majae_s[part])
    }

    results_list[[method]] <- row
  }

  # --- Column names ---
  metrics <- c("MPJPE (m)","PA-MPJPE (m)","MPJVE (m/s)","VelCorr","Energy Ratio","MAJAE (deg)")
  parts <- c("Full","Upper","Lower","Trunk")

  col_names <- c()
  for (p in parts) {
    for (m in metrics) {
      col_names <- c(col_names, paste(p, m, sep="_"))
    }
  }

  df <- as.data.frame(do.call(rbind, results_list))
  colnames(df) <- col_names
  df$Method <- rownames(df)
  rownames(df) <- NULL

  df[, c("Method", col_names)]
}


# Display the summary table
summary_df <- build_summary_table(
  gt_files,
  pred_files,
  method_names,
  energy_segments_list
)

print(summary_df)

colnames(summary_df)
```


### Article-like 

```{r select header, echo=FALSE, warning=FALSE, message=FALSE}
# Select the cols from summary_df
prepare_article_table <- function(df) {

  keep_metrics <- c("MPJPE", "MPJVE", "Energy")
  parts <- c("Full", "Upper", "Lower", "Trunk")

  selected_cols <- c()

  for (p in parts) {
    for (m in keep_metrics) {

      matches <- grep(
        paste0("^", p, "_", m, " "),   # space ensures exact metric match
        colnames(df),
        value = TRUE
      )

      selected_cols <- c(selected_cols, matches)
    }
  }

  # Safety check
  if (length(selected_cols) != 12) {
    stop(paste("Expected 12 columns, got", length(selected_cols)))
  }

  df_filtered <- df[, c("Method", selected_cols), drop=FALSE]

  return(df_filtered)
}

print(prepare_article_table(summary_df))
```

```{r article table, echo=FALSE, warning=FALSE, message=FALSE}
# Function to put 2 decimals for values between 1e-2 and 1e3, and scientific notation otherwise
format_mixed <- function(x) {
  ifelse(abs(x) >= 1e-2 & abs(x) < 1e3,
         sprintf("%.2f", x),
         formatC(x, format = "e", digits = 2))
}

# Article-like table plot
plot_article_table <- function(df) {

  df <- prepare_article_table(df)

  if (ncol(df) != 13) {
    stop("Expected 12 metric columns + Method")
  }

  # Body
  body <- df
  
  # Format numeric columns to 2 decimals
  body[,-1] <- lapply(body[,-1], format_mixed)

  # Base table (WITHOUT headers)
  g <- tableGrob(body,
                 rows = NULL,
                 cols = NULL,
                 theme = ttheme_minimal(
                   core = list(fg_params = list(cex=0.65))
                 ))

  # ---- HEADER ROW 2 (MPJPE / MPJVE / Energy) ----
  header2 <- c("Method", rep(c("MPJPE (m)","MPJVE (m/s)","Energy Ratio"), 4))

  header2_grob <- tableGrob(t(header2), rows=NULL,
                            theme = ttheme_minimal(
                              core = list(fg_params=list(fontface="bold", cex=0.7))
                            ))

  # ---- HEADER ROW 1 (Full / Upper / Lower / Trunk with spans) ----
  header1 <- gtable(widths = g$widths, heights = unit(1, "lines"))

  labels <- c("Body part", "Full", "Upper", "Lower", "Trunk")
  spans  <- c(1, 3, 3, 3, 3)

  col_start <- 1

  for (i in seq_along(labels)) {

    header1 <- gtable_add_grob(
      header1,
      textGrob(labels[i], gp=gpar(fontface="bold", cex=0.7)),
      t=1,
      l=col_start,
      r=col_start + spans[i] - 1
    )

    col_start <- col_start + spans[i]
  }

  # ---- STACK TABLE ----
  g_final <- gtable_rbind(header1, header2_grob, g)

  grid.newpage()
  grid.draw(g_final)
}

plot_article_table(summary_df)
```




